White Wine Alcohol Analysis by Subha Radhakrishnan
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(GGally)


```

```{r echo=FALSE, warning=FALSE ,fig.width=6,fig.height=3,Load_the_Data}



list.files()

wines<-read.csv("wineQualityWhites.csv")


```
Input variables (based on physicochemical tests):
1 - fixed acidity (tartaric acid - g / dm^3)
2 - volatile acidity (acetic acid - g / dm^3)
3 - citric acid (g / dm^3)
4 - residual sugar (g / dm^3)
5 - chlorides (sodium chloride - g / dm^3
6 - free sulfur dioxide (mg / dm^3)
7 - total sulfur dioxide (mg / dm^3)
8 - density (g / cm^3)
9 - pH
10 - sulphates (potassium sulphate - g / dm3)
11 - alcohol (% by volume)
Output variable (based on sensory data): 
12 - quality (score between 0 and 10)


Description of attributes:

1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)

2 - volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste

3 - citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines

4 - residual sugar: the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet

5 - chlorides: the amount of salt in the wine

6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine

7 - total sulfur dioxide: amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine

8 - density: the density of water is close to that of water depending on the percent alcohol and sugar content

9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale

10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant

11 - alcohol: the percent alcohol content of the wine

Output variable (based on sensory data): 
12 - quality (score between 0 and 10)

<h4 style="color:blue";font-size: 10pt;text-decoration: underline;>Univariate Plots Section</h4>

```{r echo=FALSE,fig.width=6,fig.height=3, warning=FALSE}

dim(wines)

names(wines)

head(wines)

summary(wines)

str(wines)

colnames(wines)[2] <- "Facidity"
colnames(wines)[3] <- "Vacidity"
colnames(wines)[4] <- "citric"
colnames(wines)[5] <- "res.Sugar"
colnames(wines)[7] <- "FSo2"
colnames(wines)[8] <- "TSo2"

wines$bound.sulfur.dioxide<- wines$TSo2-wines$FSo2

```

My goal is to associate quality to the wines based on the alcohol , density and residual sugar. In this journey I may analyse other vraiables also depending on the study of plots.
I changed the names of attributes to accomodate them on the plots.


```{r echo=FALSE,  warning=FALSE,fig.width=6,fig.height=3,Univariate_Plots}

qplot(x = quality, data = wines,color=I("Black") ,fill=I("light blue"))+
scale_x_continuous(lim = c(3 , 10), 
                breaks = seq(3 , 10, 1))+
 theme(axis.text=element_text(size=10),
        axis.title=element_text(size=12,face="bold"))
```

Quality of wine - limiting the x axis as it is clear that quality starts from 3
Quality 6 and 7 .
From the Quality chart it looks like most frequency of the wines dataset falls in the quality 6 category which can be considered as medium quality.So for my analysis I expect most of the wines falling in this medium category.I would also like to the great quality wines and what contributes to that high quality.


```{r echo=FALSE, warning=FALSE,fig.width=6,fig.height=3}

qplot(x=density, data = wines, color=I("Black"), binwidth = 0.0002) +
  scale_x_log10(lim = c(0.9871 , 1.0037), 
                breaks = seq(0.9871 , 1.0037, 0.002))+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=12,face="bold"))
```

Density of wine denotes Bimodal distribution with density high for certain groups and less for others. There is also an outlier at 1.0031
Summary gives the mean and median with minimum and maximum values. Density will help determine my analysis based on the sugar content and also come to a conclusion if sweet wine is considered to be good quality wines. 


```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE}

summary(wines$Vacidity)


qplot(x=Vacidity, data = wines, color=I("orange"),binwidth = 0.1) +
  scale_x_log10(breaks = seq(0.0800,1.1000, 0.3))+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=12,face="bold"))

```

Volatile acidity-All wines contain volatile acidity and it contributes to the overall acidity and aroma of wines.
Volatile acidity has normal distribution. It indicates that high acidity levels can decrease the taste and that is why there is always a medium balance.This may or maynot help to the analysis as taste differs for many.Good quality wines may not be tasty for me but could be considered tasty for many.


```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE}
summary(wines$alcohol)

qplot(x=alcohol, data =wines, color=I("blue"), binwidth = 0.5) +
  scale_x_continuous(breaks = seq(8,15,0.5), lim = c(8,15))+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=12,face="bold"))

```

Alcohol level distribution looks skewed. Most frequently wines have 9.5%.Mean is 10.51% of alcohol.Higher alcohol content is generally considered whisky . Hence I expect alcohol in medium range between 11 and 12. My impression is great quality wines will have alcohol content falling between 11 and 13.


```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE}
summary(wines$res.Sugar)

wine_sweet <- subset(wines, wines$res.Sugar > 40)

qplot(x=res.Sugar, data =wines, color=I("blue"), binwidth =1) +
  scale_x_continuous(breaks = seq(0.6,66,5), lim = c(0.6,66))+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=12,face="bold"))

```

Mostly white wines has less residual sugar. To make sure my analysis is right, I checked if there are wines with residual Sugar > 45 which denotes they are sweet. I could find only 1 wine with this composition. Hence I understood that most wines have less sweet and may be more alcohol!

```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE}
summary(wines$TSo2)


qplot(x=TSo2, data =wines, color=I("blue"), binwidth =9) +
  scale_x_continuous(breaks = seq(10,440,50), lim = c(10,440))+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=12,face="bold"))


```

I chose Total So2 as well to understand if it has any impact on quality of wine. The distribution is normal. Correlation with quality can indicate that high Supfur dioxide can have an offensive odour and taste
The total sulfur dioxide seems to a have a spread between 10 and 250 and exhibiting peak around 150.There seems to be a lower peak exhibited around the 20 mark.White wine seems to exhibit a total SO2 at higher levels

<h4 style="color:blue";font-size: 8pt;>Univariate Analysis</h4>

<h4 style="color:blue";font-size: 10pt;>What is the structure of your dataset?</h4>

All of the variables are numeric with quality being the only integer. There are 13 variables and 4898 oversations in the dataset.Quality is a continuous variable and free sulfur dioxide seems to be more discrete in nature.

<h4 style="color:blue";font-size: 8pt;>What is/are the main feature(s) of interest in your dataset?</h4>

My first main interest in the dataset to identify if there are any variables impacting the density which further impact quality. The dataset includes 4898 observations and most of the wine evaluation are 5 or 6 while the minimum quality is 3 and maximum 8.
Secondly I will choose the density and compare this against alcohol and 
residual sugar to understand the correlation between them and inturn understand the quality of wine.

<h4 style="color:blue";font-size: 8pt;>What other features in the dataset do you think will help support your investigation into your feature(s) of interest?</h4>

From the description of variables in the dataset,few of the variables that I felt will support investigation are:


Residual Sugar : To determinie the sweetness of a wine;
Free Sulfur Dioxide : Seems to become evident after a certain amount;
Chloride: Amount of salt in wine;
Alcochol content in wine;
density:density of water close to water based on alcohol;

<h4 style="color:blue";font-size: 8pt;>Did you create any new variables from existing variables in the dataset?</h4>

Based on the descriptions to the variable total sulfur dioxide(amount of free and bound forms of S02), I decided to create a variable that holds the bound forms of S02 as we already have free form.

wines$bound.sulfur.dioxide<- wines$total.sulfur.dioxide-wines$free.sulfur.dioxide


<h4 style="color:blue";font-size: 8pt;>Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?</h4>

I felt the positive skewed data on alcohol as unusual because my belief was that white wines contain more alcohol in it.But from the diagram it seems on a average is less than the highest alcohol content.
After I ran the ggpairs to understand the correlation, I thought it will be good to shorten the names of variables.The rest of the data were are looking good and no transformation seem to be needed expliciltly.

<h4 style="color:blue";font-size: 10pt;text-decoration: underline;>Bivariate Plots Section</h4>


```{r echo=FALSE,warning=FALSE,fig.width=14, fig.height=8, Bivariate_Plots}
wines_subset<-wines[,c(2:14)]

set.seed(2000)

theme_set(theme_minimal(20))

ggpairs(wines_subset[sample.int(nrow(wines_subset),2000),],axisLabels = "internal",upper=list(params=list(size=5)))+theme(axis.text = element_blank())
  

```

There exist good correlation between 

Positive Correlation:
residual.sugar vs. density = 0.826 (good correlation since greater than 0.5)
residual.sugar vs. total.sulfur.dioxide 0.429
fixed.sulfur.dioxide vs. total.sulfur.dioxide =0.607
alcohol vs quality = 0.436

Negative Correlation:
alcohol vs density = -0.804 (less than -0.3)
residual.sugar vs. alcohol = -0.449
density vs quality=-0.354
residual sugar vs quality= -0.117

Correlation : -0.354
```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE}
summary(wines$quality)
summary(wines$density)

ggplot(aes(y=quality,x=density),
       data = wines)+
   geom_line(aes(color='quality'),stat='summary',fun.y=median)+
     theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))+
ggtitle('Quality based on density') +
  theme(plot.title = element_text(size = 12,colour="Dark blue"))

```

From the plot above, less dense the wine, the quality is good. Now, I would need to understand how density is related to other varaiables.As density increases, perceived quality declines.Quality remains 6 if the density increases but this will be based on other dependent factors such as residual sugar or alcohol.

Correlation : 0.826
```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE}
summary(wines$res.Sugar)
summary(wines$density)

ggplot(data = wines, aes(y = density, x = res.Sugar)) +
  geom_point() +
  coord_cartesian(xlim=c(0.6,66), 
                  ylim=c(0.9871, 1.0390))+
  geom_smooth(method = 'lm')+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))

```

Residual Sugar and density go to-gether.Hence it makes sense that less the residual sugar, less the density of wine with more alcohol.

Correlation : -0.804
```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE}
summary(wines$alcohol)
summary(wines$density)

  
ggplot(data = wines, aes(y = density, x = alcohol)) +
  geom_point() +
  coord_cartesian(xlim=c(8,15), 
                  ylim=c(0.9871, 1.005))+
  geom_smooth(method = 'lm') +
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))

```


The above plot shows that less dense ,the more the amount of alcohol in the wine.
Density is directly related alcohol, in that alcohol causes change in density, it would be more prudent to say that alcohol affects both density and quality, or, density is a result of alcohol

Correlation : residual sugar vs alcohol= -0.447

Correlation : residual sugar vs quality= -0.117

```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE}

ggplot(data = wines, aes(y = res.Sugar, x = alcohol )) +
  geom_point(aes(color=quality),alpha=1/4, position = 'jitter') +
    ylim(0.6, quantile(wines$res.Sugar, 0.95)) +
    xlim(8, quantile(wines$alcohol, 0.95)) +
  geom_smooth(method = 'lm')+ theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))+
 scale_colour_gradient(low="black", high="red")


```

Less residual sugar is present on high alcohol content . Quality of wine is also high if residual sugar is less.Residual sugar should balance with acidity, alcohol, acids . It can be used to ensure a wine ages well, always evolving into a complex and interesting wine

Correlation : TSo2 vs alcohol= -0.464

Correlation : TSo2 vs quality= -0.179

```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE}


ggplot(data = wines, aes(y = TSo2, x = alcohol)) +
  geom_point() +
  coord_cartesian(xlim=c(8,15), 
                  ylim=c(9, 440))+
  geom_smooth(method = 'lm')+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))


ggplot(aes(x = quality, y = TSo2), data = wines) + 
  geom_point(aes(color='red'),alpha=1/4, position = 'jitter')+
     theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))+
  ggtitle(' Total SO2 and  Quality Relationship')+
  theme(plot.title = element_text(size = 12,colour="Dark blue"))

```

Total SO2 content decreases with Alcohol content for White wine.It's clear to see that the proper additions of SO2 are a real "treat"(quality) to people who love wine. Without it, we wouldn't have such a great diversity of high quality wines that we enjoy today.

Correlation : Chlorides vs quality= -0.25
```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE}
summary(wines$chlorides)
summary(wines$quality)


ggplot(data = wines, aes(y = chlorides, x = quality )) +
  geom_point() +
  coord_cartesian(ylim=c(0.00900,0.1), 
                  xlim=c(4,8))+
  geom_smooth(method = 'lm')+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))

```

I decided to look at chlorides and residual sugar to see how they are related since I thought they cannot exist in similar quantities.The perceived quality of  wines decline as the amount of chlorides, or salt, increase.

Correlation : Chlorides vs residual sugar= 0.0965

Correlation : Chlorides vs alcohol= -0.361


```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE}


ggplot(data = wines, aes(y = chlorides, x = res.Sugar)) +
  geom_point(position = 'jitter') +
   ylim(0.00900, quantile(wines$chlorides, 0.95)) +
    xlim(0.6, quantile(wines$res.Sugar, 0.95)) +
   geom_smooth(method = 'lm')+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))
  

ggplot(data = wines, aes(y = chlorides, x = alcohol )) +
  geom_point(position = 'jitter') +
     ylim(0.00900, quantile(wines$chlorides, 0.95)) +
    xlim(8, quantile(wines$alcohol, 0.95)) +
  geom_smooth(method = 'lm')+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))

```

The correlation of chloride and residual sugar and alcohol seems to be 
interesting as high alcohol has lower residual sugar. Between residual sugar and chloride, residual sugar is less even if chloride is high.I was hoping to see high alcohol when cloride is high but I see that alcohol is less when chloride is high.

Finally to know the quality based on alcohol:

Correlation : alcohol vs quality= 0.461

```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE}

ggplot(data = wines, aes(y = alcohol, x = quality )) +
  geom_point(position = 'jitter') +
  coord_cartesian(ylim=c(8,15), 
                  xlim=c(3,9))+
  geom_smooth(method = 'lm')+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))

```

The plot shows that alcohol alone cannot be contributed to high quality. Hence in the multivariate scetion I will use alcohol with other vraibles along with quality

<h4 style="color:blue";font-size: 6pt;>Bivariate Analysis</h4>

<h4 style="color:blue";font-size: 8pt;>Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?</h4>

With ggpair I was able to draw the correlation between a few variables which I had all plotted above.
I had no real expectations with regards to the impact of other variables on the density, there are many variables which are significantly correlated to the density of wine. 

<h4 style="color:blue";font-size: 8pt;>Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?</h4>

First one that came to mind was the relationship between Chlorides and Residual Sugar. I am not sure how to consider the observations which are on the top-left-hand corner and the ones on the bottom-right-hand corner. Would they be considered as outliers or related to the limited data available for wines with more than 4 mg/L of residual sugars OR more than 0.2 mg/L of chlorides ?

Another scatter plot with Chlorides is with Density and also with Alcohol. There seems to be a linear distribution, however few Chlorides surprisingly have high between density .995 and 1. It is not clear as to why a dozen of observations have high level of Chlorides.

<h4 style="color:blue";font-size: 8pt;>What was the strongest relationship you found?</h4>

The strongest relationship I found was 0.922 (bound SO2 and total SO2), however I decided to have a look at the second strongest relationship as bound SO2 was the added variable which was obtained through a simple substraction between total SO2 an free SO2, therefore the relationship is evident.
The second most interesting relationship was chloride over density as it came up with a correlation coefficient of 0.835. You can find below the scatterplot to illustrate this relationship

correlation coefficient of 0.835

```{r echo=FALSE,fig.width=6,fig.height=3, warning=FALSE}
summary(wines$density)
summary(wines$chlorides)

ggplot(data = wines, aes(y = chlorides, x = density)) +
  geom_point(position = 'jitter') +
      ylim(0.009, quantile(wines$chlorides, 0.95)) +
    xlim(0.98711, quantile(wines$density, 0.95))+
  geom_smooth(method = 'lm')+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))
```

White wine exhibits lower Chloride levels across Alcohol content levels and holds a steady pattern throughout

<h4 style="color:blue";font-size: 10pt;text-decoration: underline;> Multivariate Plots Section</h4>

I was able to identify the highest correlation coefficients between 
quality,density and the other variables available in our dataset.
I decided to use density and the corelated variables to understand the quality of wine.
The legend on the right indicates the color for each score.

Variables involved are:

Alcohol
density
Residual Sugar

```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE, Multivariate_Plots}

names(wines)
summary(wines$density)
ggplot(data = wines, 
       aes(y = density, x = res.Sugar, color = factor(quality))) +
   geom_point(alpha = 1/2, position = position_jitter(h = 0), size = 2) +
   coord_cartesian(ylim=c(0.9871,1.005), xlim=c(0.6,10)) +
   scale_color_brewer(type='qual') +
   xlab('res.Sugar') +
   ylab('density') + 
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))+
   ggtitle('density vs. res.Sugar correlation by Quality') +
  theme(plot.title = element_text(size = 12,colour="Dark blue"))

```

The scatter plot shows that when density and residual sugar composition is less. I see the quality is on highr side.

```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE}
ggplot(data = wines, 
       aes(x = density, y = alcohol, color = factor(quality))) +
   geom_point(alpha = 1/2, position = position_jitter(h = 0), size = 2) +
   coord_cartesian(xlim=c(min(wines$density),1.005), ylim=c(8,15)) +
   scale_color_brewer(type='qual') +
   xlab('Density') +
   ylab('Alcohol') +stat_ellipse()+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))+
   ggtitle('Density vs. Alcohol correlation by Quality') +
  theme(plot.title = element_text(size = 12,colour="Dark blue"))
```

It is clear to come to a conclusion that if density is less, alcohol will be high and inturn considered as great quality


```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE}

wines$alcoholLabel[wines$alcohol <= 9.5 ] = "Light"
wines$alcoholLabel[9.5< wines$alcohol & wines$quality <= 12] = "Mild"
wines$alcoholLabel[12< wines$alcohol ] = "Strong"
wines$alcoholLabel = as.factor(wines$alcoholLabel)

wines$rating[5 >= wines$quality ] = "Poor"
wines$rating[5< wines$quality & wines$quality < 7] = "Good"
wines$rating[7<= wines$quality ] = "Great"
wines$rating = as.factor(wines$rating)

#?position_fill

ggplot(data= wines, aes(x=as.factor(wines$quality) , fill = wines$alcoholLabel)) +geom_histogram(position="fill")+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))

```

In the above stacked bar, for the higher quality wines there is more chance that the wine has higher alcohol percentage.
There seems to be a correlation between density and alcohol . Less dense, more alcohol which also seems to be great wines .

```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE}
ggplot(data = wines, aes(x = alcohol, y = density, color = rating)) + 
  geom_point()+ scale_color_brewer() +theme(axis.text=element_text(size=10), axis.title=element_text(size=10,face="bold"))

ggplot(data = wines, aes(x = alcohol, y = res.Sugar )) + 
  facet_wrap(~rating)+geom_point(size = 3, alpha = 1/4) + ylim(4, quantile(wines$res.Sugar, 0.95)) +xlim(8, quantile(wines$alcohol, 0.95))+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))
```

<h4 style="color:blue";font-size: 6pt;> Multivariate Analysis</h4>

<h4 style="color:blue";font-size: 10pt;>Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?</h4>

According to the above scatter plot, there seems to be a relationship between alcohol and rating of the wine. most of great wines are in the right side of the plot. More specifically, if the alcohol percentage is above 11% there seems to be a good chance that we will have a good or great wine (great wine has rating 7 or above, good ones has quality above 5). If it is more than 12% the chance is even higher.

"Light": Alcohol percentage below 10%
"Mild" : Alcohol percentage higher than 10% but below 12%
"Strong": Alcohol percentage higher than 12%

The qulaity of wine is great if the alcohol content is between 12 and 14 . Based on my analysis we can that :

Less residual sugar with Alcohol between 12 and 14 with less density is 
considered great quality


<h4 style="color:blue";font-size: 10pt;>Were there any interesting or surprising interactions between features?</h4>

One of the most interesting fact that I noted is between Alcohol ,Total Sulfurdioxide ; between Free Sulfurdioxide and Alcohol ; between bound sulfurdioxide.
Between Alcohol and Total Sulfurdioxide there is a negative correlation of 
But between free sulfur dioxide and alcohol there is no meaningful relation. But the relation between  bound sulfurdioxide(the varaible generated by me) and alcohol is high. 
So I expected a positive correlation existing between Total sulfurdioxide and Alcohol. But to my surprise I found a negative relation

```{r echo=FALSE, fig.width=6,fig.height=3,warning=FALSE}

summary(wines$bound.sulfur.dioxide)
ggplot(data = wines, aes(x = alcohol, y = TSo2, color =rating )) + 
    geom_point() +
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))+
scale_color_brewer()

ggplot(data = wines, aes(x = alcohol, y = FSo2, color =rating )) + 
    geom_point()+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold") )+
scale_color_brewer()

ggplot(data = wines, aes(x = alcohol, y = bound.sulfur.dioxide, color =rating )) +  geom_point(size = 3, alpha = 1/4) + ylim(4, quantile(wines$bound.sulfur.dioxide, 0.95)) +xlim(8, quantile(wines$alcohol, 0.95)) +
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))+
scale_color_brewer()
```

Only a proportion of the SO2 added to a wine will be effective as an 
anti-oxidant. The rest will combine with other elements in the wine and cease to be useful. The part lost into the wine is said to be bound, the active part to be free.

OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

<h4 style="color:blue";font-size: 10pt;text-decoration: underline;>Final Plots and Summary</h4>

<h4 style="color:blue";font-size: 8pt;>Plot One</h4>
```{r echo=FALSE, fig.width=6,fig.height=3, warning=FALSE,Plot_One}

qplot(x=alcohol, data =wines, color=I("blue"), binwidth = 0.5) +
  scale_x_continuous(breaks = seq(8,15,0.5), lim = c(8,15))+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))+
  ggtitle('Alcohol spread in wine') +
  theme(plot.title = element_text(size = 12,colour="Dark blue"))
```

<h4 style="color:blue";font-size: 8pt;>Description One</h4>
My first plot was to look at the distribution of the wine scores in order to understand the scope of the dataset and  I was surprised by the distribution of the alcohol percentage which I sought to be higher between 13-14.The distribution of alcohol on this dataset was mostly concentrated into the 9-10 degrees

<h4 style="color:blue";font-size: 8pt;>Plot Two</h4>
```{r echo=FALSE,fig.width=6,fig.height=3, warning=FALSE, Plot_Two}
ggplot(data = wines, aes(y = density, x = alcohol)) +
  geom_point() +
  coord_cartesian(xlim=c(8,15), 
                  ylim=c(0.9871, 1.005))+
  geom_smooth(method = 'lm') +
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))+
  ggtitle('Density vs Alcohol') +
  theme(plot.title = element_text(size = 12,colour="Dark blue"))
```

<h4 style="color:blue";font-size: 8pt;>Description Two</h4>
When looking at Bivariates plots and the GGpair output, it was clear that quality had the most significant impact  with a negative correlation on alcohol,density and residual sugar. The higher the range of alcohol in the wine, the higher the quality is likely to be. In order to confirm this observation, you can have a look at the histogram above and see the distribution of the percentage of alcohol within wines grouped by scores.

<h4 style="color:blue";font-size: 8pt;>Plot Three</h4>
```{r echo=FALSE,fig.width=6,fig.height=3, warning=FALSE, Plot_Three}

ggplot(data = wines, aes(x = alcohol, y = res.Sugar,colour=factor(quality),alpha=Complexity )) + 
  facet_wrap(~rating)+geom_point(size = 3, alpha = 1/4) + ylim(4, quantile(wines$res.Sugar, 0.95)) +xlim(8, quantile(wines$alcohol, 0.95))+
   theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold"))+
  ggtitle('Quality based on Alcohol and residual sugar') +
  theme(plot.title = element_text(size = 12,colour="Dark blue"))
```

<h4 style="color:blue";font-size: 8pt;>Description Three</h4>
From the Multivariate plot section I found a good correlation between 
alcohol,density,residual sugar by which the quality could also be determined.
Hence it became clear from the dataset that Less dense and less residual sugar and alcohol in the range 10-14 are considered to be of great quality
------

<h4 style="color:blue";font-size: 10pt;text-decoration: underline;>Reflection</h4>

Assuming we can trust the data, there are some interesting illuminations. More alcohol is associated with higher quality wines, density can fairly well predicted by the chemical information in the data set.
One important factor that contributes to Alcohol percentage is the remaining sugar in wine after fermentation so that the more sugar left after fermentation, the less the percentage of alcohol will be in the wine.


Analyzing the data the following conclusion:

When alcohol percentage decreases, density grows.
Total sulfur dioxide and level of residual sugar are positively correlated. 
White wine density and residual sugar level have positive correlation.
Alcohol level of  wine decreases with the growth of residual sugar level.
Wine with high alcohol percentage has quality level 7, wine with less alcohol percentage is quality level 5.
Mostly frequent quality levels of white wine are 5 and 6.

Success Factor with dataset:
I can still judge a wine quality (to some extent) with very technical variables without the need to actually taste it. It must provide the wine maker some sort of guidance towards how to produce a better wine. From the highest correlation coefficient, I can see that Alcohol plays a big part, therefore the wine makers must try to find techniques in order to increase the alcohol degree towards 14 and also decrease the volatile acidity (negative correlation). 

Though I was able to judge the impact analysis on quality of wine, it would have been complete if more data was available. The biggest struggle was to analyse every correlated varaible ass the variables were all very close.
I would be able to produce a better model as there's a lot of impactful data that were not available in this dataset such as :

Grape Types

Year

Region


